### 2022-12-17 Figuring_Q3.Rmd
Starting to work on code producing pdf report

---
title: "Untitled"
output: html_document
date: "2022-12-14"
---
```{r}
# Part b
calcDirichletPDF <- function(x,a){
  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

#Part c
simDirichletRV <- function(a){
  #Drawing from gamma distribution
  U <- rgamma(a,1)
  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
  V <- U/sum(U)

  return(V)
}

###Testing 
#Initial values
#finding prob of sequence under current pt
X <- read.table("q3Weather.txt",header = F)
X <- X$V1
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)
n <- 500000
burnin <- 5

  #Converting X into a numeric vector to access entries of PT
  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
  init.state <- X[1:(length(X)-1)]
  trans.state <- X[2:length(X)]

  #initializing algorithm
  V <- array(NA,dim=c(3,3,n+burnin+1))
  V[,,1] <- initStateProbs
  est.state.prob <- array(NA,dim=c(1,3,n+burnin+1)) 
  est.state.prob[,,1] <-  initStateProbs

proposal <- matrix(NA,ncol=3,nrow=3)
#starting loop
for (j in 2:(n+burnin+1)){

    est.state.prob[,,j] <- est.state.prob[,,j-1]%*%V[,,j-1]
    #Generating proposal value for each row of transition matrix
    for (i in 1:length(initStateProbs)){
      proposal[i,] <- simDirichletRV(20*V[i,,j-1])
    }
     #calculating alpha

      #Generating posterior using equation (12) for proposal and old value of PT
      actualp.prev <- actualp.prop <- NULL
      for (i in 1:length(init.state)){
        actualp.prev[i] <- V[init.state[i],trans.state[i],j-1]
        actualp.prop[i] <- proposal[init.state[i],trans.state[i]]
      }
      #density has some constant as proportional but due to fraction constant cancels out so dont
      #need to worry about it
      true.density.prev <- prod(actualp.prev)*initStateProbs[X[1]]
      true.density.prop <-prod(actualp.prop)*initStateProbs[X[1]]
      #alpha <- true.density.prop/true.density.prev

      #Doing shady stuff (assuming that px` or previous is the x vector :P)
      #Calculating pdf under proposal using x as the probability to be in each state 
       prop.density.prop <- exp(calcDirichletPDF(x=proposal[2,],a=est.state.prob[,,j-1])+
                                  calcDirichletPDF(x=proposal[1,],a=est.state.prob[,,j-1])+calcDirichletPDF(x=proposal[3,],a=est.state.prob[,,j-1]))
      # #same but under previous accepted value
       prop.density.prev <- exp(calcDirichletPDF(x=V[2,,j-1],a=est.state.prob[,,j-1])+
                                 calcDirichletPDF(x=V[1,,j-1],a=est.state.prob[,,j-1])+calcDirichletPDF(x=V[3,,j-1],a=est.state.prob[,,j-1]))

      # alpha <- (true.density.prop*prop.density.prev)/(true.density.prev*prop.density.prop)
      alpha <-  prop.density.prop/prop.density.prev

      if(min(1,alpha)>=runif(1)){
        V[,,j] <- proposal
      }else{
        V[,,j] <- V[,,j-1]
      }
}

# A <- array(NA,dim=c(3,3,n+burnin+1))
# A[,,1] <- matrix(c(1,2,3,4,5,6,7,8,9),ncol=3,nrow=3,byrow = T)
# #1st index is row
# A[1,,1]
# #2nd index is column
# A[,1,1]

plot(V[1,2,],type="l") #No convegerence :(
```

```{r}
## Question 3

#start by doing b and c 
#then do d awhile figuring out a (the algorithm part)

#Part b
calcDirichletPDF <- function(x,a){
  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

#Part c
simDirichletRV <- function(a){
  #Drawing from gamma distribution
  U <- rgamma(a,1)
  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
  V <- U/sum(U)

  return(V)
}

#Part d
q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
  #X is vector of weather for each observed day (d long if recorded weather for d days)
  #Converting X into a numeric vector to access entries of PT
  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
  init.state <- X[1:(length(X)-1)]
  trans.state <- X[2:length(X)]

  #initializing algorithm
  V <- array(NA,dim=c(3,3,n+burnin+1))
  V[,,1] <- initStateProbs

  for (j in 2:(n+burnin+1)){
    proposal <- matrix(NA,ncol=3,nrow=3)
    #Generating proposal value for each row of transition matrix
    for (i in 1:length(initPT)){
      proposal[i,] <- simDirichletRV(20*V[,i,j-1])
    }

      #Generating posterior using equation (12) for proposal and old value of PT
      prob.trans.prev <- NULL
      prob.trans.curr <- NULL
      for (i in 1:length(init.state)){
        prob.trans.prev[i] <- V[init.state[i],trans.state[i],j-1]
        prob.trans.curr[i] <- proposal[trans.state[i],init.state[i]]
      }
      #density has some constant as proportional but due to fraction constant cancels out so dont
      #need to worry about it
      density.prev <- prod(prob.trans.prev)*initStateProbs[X[1]]
      density.curr <-prod(prob.trans.curr)*initStateProbs[X[1]]

      #alpha simplier as q is symmetric (Big boi assumption may not be able to do this)
      alpha <- min(1,density.curr/density.prev)

      #Accept or Reject
      if (runif(1) <= alpha){
        V[,,j] <- proposal
      } else{
        V[,,j] <- V[,,j-1]
      }
  }

  return(V[,,(burnin+1:n+burnin+1)])

}
#finding prob of sequence under current pt
X <- read.table("q3Weather.txt",header = F)
X <- X$V1

#Part e
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)

q3.mcmc(X,initStateProbs,initPT,10000,100000)
```