### 2022-12-21 draft_report.Rmd
Adding in draft of explanation of Q2

---
title: "STatistical Computing (MATH6173) Coursework 2"
author: "Student ID: 34273638"
geometry: margin = 2cm
output:
  pdf_document: default
header-includes:
  - \usepackage{wrapfig}
  - \usepackage{lipsum}
  - \usepackage{setspace}
  - \usepackage{titlesec}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{mathtools}
  - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
---

```{r, include=F}
#Setting up enviroment
setwd("~/Desktop/Statistical Computing/CW/SC-CW2")
```

# Question 1

## Part a: (add graphs to help in explanation)

\begin{equation}
\pi(x) =
\begin{cases}
 \frac{as^{a}}{2x^{a+1}}\, ,& s<x, \\
 \frac{1}{2b}\, ,& s-b \leq x \leq s, \\
 0  ,& \text{otherwise,}
 \end{cases}  
 \end{equation}

\begin{equation}
g(x) = \frac{\lambda}{2}exp(-\lambda|x-b|)
\end{equation}

For acceptance-rejection sampling we need to find constant $M>0$ such that $\pi(x) \leq Mg(x)$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density. This constant M determines the acceptance probability and hence the efficiency of the algorithm. With the acceptance probability (denoted $\tau$) equal to $\frac{1}{M}$. For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have the higher acceptance probability. To find M for each $\lambda$ value I split the real domain of x into 3 regions (see plot 1) corresponding to the different PARTS of $\pi$ and study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith interval. Then $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of x.

For the first region ($-\infty, s-b$) $\pi(x)=0$ thus we can let $M_{1}$ be some arbitrarily small value greater than 0 say $\epsilon$ and this will satisfy $\pi(x) \leq M_{1}g(x)$ in the region specified.
For the second region [$s-b,s$], $\pi(x)=\frac{1}{2b}$. Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all x in the region.
For the third region ($s,\infty$), $\pi(x)=\frac{as^{a}}{2x^{a+1}}$. I note that $\pi(x)$ is a decreasing function in this region (see plot 1) and initially $g(x)$ is a smaller but increasing function (see plot 1), until it reaches its maximum (larger than $\pi(x)$ at this value of x) where it then decreases for all larger x but always is larger than $\pi(x)$. Thus setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all x in this region.

Subbing in the values of a=2, b=3 and s=1.25 given in the question the value of M for $\lambda=1$ is $M=max(\epsilon,\frac{e^{4.75}}{3},1.6e^{1.75}) \approx 38.53$ and for $\lambda=0.5$ is $M=max(\epsilon,\frac{e^{2.375}}{1.5},3.2e^{0.875})\approx 7.7$. Thus as M is smaller for $\lambda=0.5$. Thus, out of these 2 values for lambda 0.5 is my chosen value as it leads to an algorithm with a higher efficiency.

```{r,echo=F,fig.align='center'}
#a
#Creating plots to help explain

gx.1 <- function(x){
  gx <- 0.5*exp(-abs(x-3))
  return(gx)
}

gx.2 <- function(x){
  gx <- 0.25*exp(-0.5*abs(x-3))
  return(gx)
}

target.dist<- function(x){
  pi <-NULL
  for (i in 1:length(x)){
  if (x[i]>1.25){
    pi[i] <- (2*1.25^(2))/(2*x[i]^(3))
  } else if (x[i] >= -1.75){
    pi[i] <- 1/6
  }else{
    pi[i] <- 0
  }
  }
  return(pi)
}

x<- seq(from=-4, to=10, by=0.01)

plot(x=x,y=target.dist(x),
     type="l",
     ylab = "density",
     main="Plot 1: Comparing target and proposal pdf's",
     sub="Note: used a=2,b=3,s=1.25 with the green line seperating the regions of the graph")
curve(gx.1(x),add=T,col="red")
curve(gx.2,add=T,col="blue")
abline(v=-1.75,lty=2,col="green")
abline(v=1.25,lty=2,col="green")
text(x=-3,y=0.4,labels="Region 1")
text(x=0,y=0.4,labels="Region 2")
text(x=6,y=0.4,labels="Region 3")
legend("topright", lty = c(1,1,1), cex = 0.6, col = c("black","red","blue"),
legend = c("Target distribution pdf","g(x) pdf with lambda=1","g(x) pdf with lambda=0.5"))
```

## Part b:
To implement A-R sampling using the proposal density with $\lambda=0.5$ to obtain samples from $\pi(x)$ I use the following algorithm:

To sample from $g(x)$ I use the inverse transformation technique by drawing u from a uniform distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$ else set $x=b-\frac{2-2u}{\lambda}$. Then after obtaining x a draw from $g(x)$ perform A-R sampling by drawing y from a uniform(0,1) distribution and if $s < x$ then checking if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{as^{a}}{2x^{a+1}}\frac{2}{M\lambda exp(-\lambda|x-b|)}$ else if $s-b <x$ then checking if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{1}{2b}\frac{2}{M\lambda exp(-\lambda|x-b|)}$ if either of these checks are true then x is a draw from the target distribution $\pi(x)$ and we save this value, if these checks are not true or $x<s-b$ then x is not a draw from $\pi(x)$ and we throw away x. I repeat this process until the desired number of samples from $\pi(x)$ has been attained.

## Part d:
```{r,include=F}
#c
q1.ar <- function(n,a,b,s,lambda,M){

    X <- NULL
    for (i in 1:n){
      repeat{
        #Drawing x from the proposal distibution 
        u <- runif(1)
        if (u<0.5){
          x <- b+log(2*u)/lambda
        }else {
          x <- b-log(2-2*u)/lambda
        }
        #Performing A-R sampling
        y <- runif(1)
        if (s < x){
          if(y <= (a*s^a)/(2*x^(a+1))/(M*lambda*exp(-lambda*abs(x-b))/2)){
            X[i]<-x
            break
          }
        } else if(s-b<x){
          if(y <= 1/(2*b)/(M*lambda*exp(-lambda*abs(x-b))/2)){
            X[i]<-x
            break
          }
        }
      }
    }
    return(X)
}

#d
set.seed(2022)
ar.sample <- q1.ar(5000,2,3,1.25,0.5,M=3.2*exp(0.875))
#Creating function to create actual pdf of target distribution
target.dist <- function(x,a,b,s){
  n<-length(x)
  y<-rep(0,n)
  for (i in 1:n){

  if (x[i]<s-b){
    y[i] <- 0
  }else if (x[i]>s){
    y[i] <- (a*s^a)/(2*x[i]^(a+1))
  } else {
    y[i] <- 1/(2*b)
  }
  }
  return(y)
}
```

```{r,echo=F,fig.align='center'}
#Creating curve of target distribution
hist(ar.sample,nclass=100,prob = T,
     xlab = "Values generated by AR sampling",
     main = "Simulated and actual pdf of target distribution");
curve(target.dist(x,2,3,1.25),from=-4,to=20,n=1000,add=T, lwd= 1, lty= 1,col="red");
legend("topright", lty = c(1), cex = 0.6, col = c("red"),
legend = c("Actual pdf"))
```

```{r,include=F}
#e
#Using the plain Monte Carlo estimator of I (I_{n})

#Function to calculate the kth root of kth moment using plain Monte Carlo estimator
k.root.k.moment <- function(sample,n,k){
  output <- (1/n*sum(sample^k))^(1/k)
  return(output)
}

#k=2
round(k.root.k.moment(ar.sample,5000,2),3)
#2nd root of 2nd moment is 2.204 (3 d.p.)

#k=3
round(k.root.k.moment(ar.sample,5000,3),3)
#3rd root of 3rd moment is 3.350 (3 d.p.)
```

# Question 2

## Part a
For the continuous-time random walk conditioned on $U_{0}=u_{0}$
\begin{equation}
U_{i} \sim \text{Normal}(\mu=U_{i-1},\sigma^{2}=t_{i}-t_{i-1})
\end{equation}

First by transforming $a$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$ $p_{tail}(a)=Pr(X>a|X_{0}=x_{0})  \equiv Pr(U>b|U_{0}=u_{0}=p_{tail}(b))$, thus we can work with the sequence $U_{i}$ to answer equation (6). Now from equation (3):
\begin{equation*}
Pr(U_{i}>b_{i})=Pr((\sqrt{t_{i}-t_{i-1}})Z+u_{i-1}>b_{i}))=Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right) \hspace{2em}  
\end{equation*}
For Z the standard normal and $u_{i-1}$ a sample from $U_{i-1}$
Therefore, to not waste any samples let $q_{i}$ the proposal the ith observation in the sequence $U_{i}$ be an exponential distribution shifted by $s_{i}$ where:

\begin{equation*}
 s_{i}=\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} 
\end{equation*}

As the samples $w=\{w_{1},...,w_{d}\}$ drawn from these proposals $q=\{q_{1},....,q_{d}\}$ will satisfy $w_{i}>b_{i},\hspace{1em} i \in \{1,...,d\}$ thus all values can be used to calculate $p_{tail}(b)=p_{tail}(a)$.

To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$). Thus, each iteration will need to generate these $u_{i}$ values using equation (3) the initial value $u_{0}$ and times $t_{i}i \in \{0,...,d\}$ to generate a realization $u$ of the random walk sequence $U$ and storing these in a vector to access to generate a sample from the proposal distribution.

## Part b

The algorithm will start by applying the logit transform to a and then using these transformed values b to work with values in sequence

First by transforming $\boldsymbol{a}$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$ $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b}))$, thus we can work with the sequence $U_{i}$ to answer equation (6). Therefore start the algorithm by transforming inputs into proportions (if given as percentages) and calculating $\boldsymbol{b}$.

Next we must simulate the random walk $U_{i}$ using equation 3 and storing these realizations $u_{i}$ in a vector.
Using these values and the proposals $q_{i}$ in part a generate $v_{i}$ a value for each step in sequence $U_{i}$
Calculate weights $w_{i}$ for each $v_{i}$ by formula $w_{i}=\frac{\pi_{i}(v_{i})}{Q_{i}(v_{i})})$ where $\pi_{i}$ is the pdf of $U_{i}$ and $Q_{i}$ is the density of the proposal distribution $q_{i}$
Then to calculate equation (6) for this iteration find the product of the weights as (put in equation with integral and how sampled pop is that integral)
Repeat this process until have desired n values and then calculate importance sampling estimator by finding mean of these values.